---
title: "CellChat_Figures"
author: "Kate Griffin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(CellChat)
library(patchwork)
library(Seurat)
library(ComplexHeatmap)

# options(future.globals.maxSize = 7000 * 1024^2)
# future::plan("multisession", workers = 4) # do parallel
```

# Load Data
merged data sets
```{r}
cellchat2 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_merged_PBS2_NP2.rds")
cellchat7 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_merged_PBS7_NP7.rds")
cellchat_PBS <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_merged_D0_PBS7.rds")
cellchat_NP <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_merged_D0_NP7.rds")
```

single data sets
```{r}
cellchat_D0 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_PBS_D0.rds")
cellchat_PBS2 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_PBS_D2.rds")
cellchat_NP2 <-readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_NP_D2.rds")
cellchat_PBS7 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_PBS_D7.rds")
cellchat_NP7 <- readRDS(file = "~/SCIFlexSeq/Data/Cellchat/cellchat_NP_D7.rds")

```

```{r}
object.list2 <- list(PBS = cellchat_PBS2, NP = cellchat_NP2)
object.list7 <- list(PBS = cellchat_PBS7, NP = cellchat_NP7)
object.list_pbs <- list(D0 = cellchat_D0, D7_PBS = cellchat_PBS7)
object.list_np <- list(D0 = cellchat_D0, D7_NP = cellchat_NP7)

```



# D0 to D7 Signaling Heatmaps
```{r d0 to d7 PBS heatmap}
gg1 <- netVisual_heatmap(cellchat_PBS)
#> Do heatmap based on a merged object
gg2 <- netVisual_heatmap(cellchat_PBS, measure = "weight")
#> Do heatmap based on a merged object
gg1 + gg2
```
```{r d0 to d7 NP heatmap}
gg3 <- netVisual_heatmap(cellchat_NP)
#> Do heatmap based on a merged object
gg4 <- netVisual_heatmap(cellchat_NP, measure = "weight")
#> Do heatmap based on a merged object
gg3 + gg4
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D0toD7_PBS_NP_Heatmap.pdf", 
    width = 8, 
    height = 5)
gg1 + gg3
dev.off()
```


# Monocytes and Neutrophils outgoing pathways
```{r}
# netAnalysis_river(cellchat_NP7, pattern = "outgoing")
# PBS is squares
h1 = netAnalysis_dot(cellchat_NP2, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 21)
h2 = netAnalysis_dot(cellchat_NP7, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 21)
h3 = netAnalysis_dot(cellchat_PBS2, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 22)
h4 = netAnalysis_dot(cellchat_PBS7, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 22)

patchwork::wrap_plots(h3, h1, nrow=2)
patchwork::wrap_plots(h4,h2, nrow=2)

# dev.off()
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_Outgoing_Signal_D2_Dot.pdf", 
    width = 16, 
    height = 4)
patchwork::wrap_plots(h3, h1, nrow=2)
dev.off()
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_Outgoing_Signal_D7_Dot.pdf", 
    width = 16, 
    height = 4)
patchwork::wrap_plots(h4,h2, nrow=2)
dev.off()
```

# Heatmap all signaling
```{r}
library(ComplexHeatmap)
 # D2 outgoing signal pattern heatmaps PBS vs NP
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list2[[i]]@netP$pathways, object.list2[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list2[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list2)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list2[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list2)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```
```{r}
 # D7 outgoing signal pattern heatmaps PBS vs NP
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list7[[i]]@netP$pathways, object.list7[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list7[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list7)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list7[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list7)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```
# Heatmap Specific Pathways
```{r}
pathways.show <- c("VWF", "Demosterol", "LIGHT", "PERIOSTIN", "PARs", "SEMA5", "MMP", 
                   "L1CAM", "PVR", "ADGRB", "CDH5", "VEGF", "JAM", "APP", "TNF", "LAMININ", "COLLAGEN"
)

# combining all the identified signaling pathways from different datasets 
# pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)

object.list <- object.list7
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/CellChat_D7_signaling_patterns_specific_paths_all.pdf",width = 8, height = 6)
i=1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)

draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

dev.off()
```

```{r}
pathways.show <- c("VWF", "Demosterol", "LIGHT", "PERIOSTIN", "PARs", "SEMA5", "MMP", 
                   "L1CAM", "PVR", "ADGRB", "CDH5", "VEGF", "JAM", "APP", "TNF", "LAMININ", "COLLAGEN"
)

# combining all the identified signaling pathways from different datasets 
# pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)

object.list <- object.list2
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/CellChat_D2_signaling_patterns_specific_paths_all.pdf",width = 8, height = 6)
i=1
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)

draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

dev.off()
```

# 2D Plot Source and Targets
```{r}
num.link <- sapply(object.list2, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list2)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list2[[i]], title = names(object.list2)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg)
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Incoming_Outgoing_Scatter_D2.pdf", 
    height = 5, width = 12)
patchwork::wrap_plots(plots = gg)
dev.off()
```

```{r}
num.link <- sapply(object.list7, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list7)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list7[[i]], title = names(object.list7)[i], weight.MinMax = weight.MinMax)
}
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
#> Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
patchwork::wrap_plots(plots = gg)
```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Incoming_Outgoing_Scatter_D7.pdf", 
    height = 5, width = 12)
patchwork::wrap_plots(plots = gg)
dev.off()
```

# DotPlots
## M+N to Endo
```{r}
levels(cellchat2@meta$phenotype)
levels(cellchat7@meta$phenotype)
```


```{r}
# from: mon (3) neut (7)
# to: endo (2)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1 <- netVisual_bubble(cellchat2, sources.use = c(3,7), targets.use = c(2),  comparison = c(1, 2), angle.x = 45)
gg2 <- netVisual_bubble(cellchat7, sources.use = c(3,7), targets.use = c(2),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
patchwork::wrap_plots(plots = list(gg1,gg2))

# dev.off()
```


```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_to_Endo_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```



## M+N to microglia
```{r}
# from: mon (3) neut (7)
# to: endo (2)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Micro_LR_Dotplot.pdf",width = 8, height = 14)

gg1 <- netVisual_bubble(cellchat2, sources.use = c(3,7), targets.use = c(1),  comparison = c(1, 2), angle.x = 45)
gg2<- netVisual_bubble(cellchat7, sources.use = c(3,7), targets.use = c(1),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_to_Micro_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```


## M+N to stromal
```{r}
# from: mon (3) neut (7)
# to: stromal (5), pericyte (6), fibroblast (13)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1<- netVisual_bubble(cellchat2, sources.use = c(3,7), targets.use = c(5),  comparison = c(1, 2), angle.x = 45)
gg2<-netVisual_bubble(cellchat7, sources.use = c(3,7), targets.use = c(5),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_to_Stro_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

## M+N to fibroblast
```{r}
# from: mon (3) neut (7)
# to: stromal (5), pericyte (6), fibroblast (10)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1<- netVisual_bubble(cellchat2, sources.use = c(3,7), targets.use = c(10),  comparison = c(1, 2), angle.x = 45)
gg2<-netVisual_bubble(cellchat7, sources.use = c(3,7), targets.use = c(10),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_to_Fib_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

## M+N to pericyte
```{r}
# from: mon (3) neut (7)
# to: stromal (5), pericyte (6), fibroblast (13)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1<- netVisual_bubble(cellchat2, sources.use = c(3,7), targets.use = c(6),  comparison = c(1, 2), angle.x = 45)
gg2<- netVisual_bubble(cellchat7, sources.use = c(3,7), targets.use = c(6),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/MonNeut_to_Peri_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

## to M
```{r}
# from: mon (3) neut (7)
# to: stromal (5), pericyte (6), fibroblast (13)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1<- netVisual_bubble(cellchat2, #sources.use = c(3,7), 
                       targets.use = c(3),  comparison = c(1, 2), angle.x = 45)
gg2<- netVisual_bubble(cellchat7, #sources.use = c(3,7), 
                       targets.use = c(3),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/All_to_Mon_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

## to N
```{r}
# from: mon (3) neut (7)
# to: stromal (5), pericyte (6), fibroblast (13)
# cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_MonNeut_to_Endo_LR_Dotplot.pdf",width = 8, height = 14)

gg1<- netVisual_bubble(cellchat2, #sources.use = c(3,7), 
                       targets.use = c(7),  comparison = c(1, 2), angle.x = 45)
gg2<- netVisual_bubble(cellchat7, #sources.use = c(3,7), 
                       targets.use = c(7),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
# dev.off()
patchwork::wrap_plots(plots = list(gg1,gg2))

```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/All_to_Neut_Dot.pdf")
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

## Outgoing from Endo

```{r}
# from: mon (3) neut (7)
# to: endo (2)
cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_Endo_Outgoing_LR_Dotplot.pdf",width = 10, height = 18)

gg1 <- netVisual_bubble(cellchat2, sources.use = c(2), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
gg2 <- netVisual_bubble(cellchat7, sources.use = c(2), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
patchwork::wrap_plots(plots = list(gg1,gg2))

dev.off()
```

## Outgoing from Micro
```{r}
# from: mon (3) neut (7)
# to: endo (2)
cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_Micro_Outgoing_LR_Dotplot.pdf",width = 10, height = 18)

gg1 <- netVisual_bubble(cellchat2, sources.use = c(1), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
gg2 <- netVisual_bubble(cellchat7, sources.use = c(1), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
patchwork::wrap_plots(plots = list(gg1,gg2))

dev.off()
```

## Outgoing from Fibro 
```{r}
# from: mon (3) neut (7)
# # to: endo (2)
cairo_pdf(file = "~/SCIFlexSeq/Figures/Cellchat/D2_Stro_Outgoing_LR_Dotplot.pdf",width = 12, height = 30)

gg1 <- netVisual_bubble(cellchat2, sources.use = c(5,10), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
gg2 <- netVisual_bubble(cellchat7, sources.use = c(5,10), targets.use = c(1,2,3,5,7,10),  comparison = c(1, 2), angle.x = 45)
#> Comparing communications on a merged object
patchwork::wrap_plots(plots = list(gg1,gg2))

dev.off()
```

# Scatter Plots 
11.65 x 4.17 inches
## M+N
```{r}
gg1 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Mon", title = "Signaling Changes in Monocytes, D2")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg2 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Mon", title = "Signaling Changes in Monocytes, D7")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1

pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Monocyte_Scatter.pdf", width = 11.65, height = 4.17)
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```
```{r}
gg3 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Neut", title = "Signaling Changes in Neutrophils, D2")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg4 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Neut", title = "Signaling Changes in Neutrophils, D7")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Neutrophil_Scatter.pdf", width =11.8, height = 3.8)
patchwork::wrap_plots(plots = list(gg3,gg4))
dev.off()
```
## Endo + Micro
```{r}
gg1 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Endo")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg2 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Endo")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Endothelial_Scatter.pdf", width =11.5, height = 3.59)
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

```{r}
gg3 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Micro")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg4 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Micro")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Microglia_Scatter.pdf", width = 11.88, height = 3.59)
patchwork::wrap_plots(plots = list(gg3,gg4))
dev.off()
# patchwork::wrap_plots(plots = list(gg3,gg4))
```

## Fib + Stro
```{r}
gg1 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Fib")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg2 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Fib")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
# patchwork::wrap_plots(plots = list(gg1,gg2))
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Fibroblast_Scatter.pdf", width =11.88, height = 3.59)
patchwork::wrap_plots(plots = list(gg1,gg2))
dev.off()
```

```{r}
gg3 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Stro")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg4 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Stro")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
# patchwork::wrap_plots(plots = list(gg3,gg4))
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Stromal_Scatter.pdf", width = 11.88, height = 3.59)
patchwork::wrap_plots(plots = list(gg3,gg4))
dev.off()

```
## Peri
```{r}
gg3 <- netAnalysis_signalingChanges_scatter(cellchat2, idents.use = "Peri")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0
#> The following `from` values were not present in `x`: 0, -1
gg4 <- netAnalysis_signalingChanges_scatter(cellchat7, idents.use = "Peri")
#> Visualizing differential outgoing and incoming signaling changes from NL to LS
#> The following `from` values were not present in `x`: 0, 2
#> The following `from` values were not present in `x`: 0, -1
# patchwork::wrap_plots(plots = list(gg3,gg4))
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Pericyte_Scatter.pdf", width = 11.88, height = 3.59)
patchwork::wrap_plots(plots = list(gg3,gg4))
dev.off()
```
# List all active pathways 
```{r}
# cellchat_NP2@netP$pathways
unique(unique(cellchat_NP2@netP$pathways,cellchat_PBS2@netP$pathways),
       unique(cellchat_NP7@netP$pathways, cellchat_PBS7@netP$pathways))

intersect(cellchat_NP2@netP$pathways,cellchat_NP7@netP$pathways)
intersect(cellchat_PBS2@netP$pathways,cellchat_PBS7@netP$pathways)

setdiff(cellchat_NP2@netP$pathways,cellchat_PBS2@netP$pathways)
setdiff(cellchat_NP7@netP$pathways,cellchat_PBS7@netP$pathways)
```

# Check which pathways are different
```{r}
# day 2
object.list <- object.list2
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```
```{r}
# day 7
object.list <- object.list7
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```


# ECM Pathways
The primary signaling pathways involved in ECM deposition are primarily mediated through integrin receptors, which activate pathways like Focal Adhesion Kinase (FAK)/Src, MAPK (Mitogen Activated Protein Kinase), and PI3K/Akt upon binding to ECM components like collagen, fibronectin, and laminin, leading to downstream effects like cell proliferation, migration, and ultimately, ECM production. 
Key points about ECM deposition signaling:
Integrin-mediated signaling:
Integrins act as the primary receptor for ECM molecules, initiating signaling cascades when engaged with their specific ligands. 
FAK/Src pathway:
Upon integrin activation, FAK and Src tyrosine kinases become phosphorylated, triggering downstream signaling events crucial for cell adhesion and migration. 
MAPK pathway:
Activated by integrin signaling, the MAPK cascade (including ERK1/2) regulates gene expression related to cell proliferation and ECM production. 
PI3K/Akt pathway:
This pathway is also activated by integrin signaling and contributes to cell survival, proliferation, and ECM deposition. 
Rho/ROCK pathway:
Integrin-mediated activation of Rho GTPase can lead to Rho-associated kinase (ROCK) activation, impacting cytoskeletal dynamics and cell motility, which are important for ECM remodeling. 
Other important factors in ECM deposition signaling:
Growth factors:
Various growth factors like TGF-β can activate specific signaling pathways to stimulate ECM production. 
Mechanical cues:
The stiffness of the ECM can influence cell behavior and ECM deposition through mechanotransduction pathways. 
Proteases:
Enzymes like matrix metalloproteinases (MMPs) play a crucial role in ECM remodeling by degrading existing ECM components, facilitating new deposition. 

GO:BP
GO[["genesets"]][["Extracellular Matrix Assembly"]]
 [1] "Agt"     "Antxr1"  "Atp7a"   "Clasp1"  "Clasp2"  "Col1a2"  "Col3a1"  "Dag1"    "Efemp2"  "Emilin1" "Fbln5"   "Fkbp10"  "Gas6"    "Gpm6b"   "Hapln2" 
[16] "Has1"    "Has2"    "Has3"    "Ihh"     "Lamb1"   "Lamb2"   "Lamb3"   "Lox"     "Ltbp3"   "Mfap4"   "Myh11"   "Notch1"  "Ntn4"    "Ntng1"   "Ntng2"  
[31] "Phldb1"  "Phldb2"  "Plod3"   "Pxdn"    "Qsox1"   "Ramp2"   "Rgcc"    "Smad3"   "Smpd3"   "Sox9"    "Tgfb1"   "Thsd4"   "Tie1"    "Tnxb" 
```{r ECM pathways}
pathways.show <- c("LAMININ", "COLLAGEN", "FN1", "MMP", "GALECTIN", "ICAM")

# day2
object.list <- object.list2
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

# day 7
object.list <- object.list7
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))


```


# Angiogensis Pathways
GO[["genesets"]][["Angiogenesis Involved In Wound Healing"]]
 [1] "Adipor2"  "Alox5"    "B4galt1"  "Cd34"     "Cx3cl1"   "Cxcr4"    "Dag1"     "Foxc2"    "Gata2"    "Gpr4"     "Gpx1"     "Hpse"     "Itgb3"   
[14] "Kdr"      "Mcam"     "Ndnf"     "Pik3cb"   "Prcp"     "Serpine1" "Slc12a2"  "Smoc2"    "Srf"      "Tafa5"    "Tnf"      "Tnfaip3"  "Vegfa"   
[27] "Xbp1"   
GO[["genesets"]][["Angiotensin Activated Signaling Pathway"]]
 [1] "Actn2"  "Agtr1a" "Agtr2"  "Agtrap" "Ahcyl1" "Car2"   "Camk2a" "Cav1"   "Mas1"   "Rock1"  "Rock2"  "Src"   
```{r angiogenesis pathways}
pathways.show <- c("THBS", "ESAM", "CXCL", "VEGFA", "PECAM1", "SEMA3")

object.list <- object.list2
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))


object.list <- object.list7
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```

# Inflammatory Pathways
GO[["genesets"]][["Acute Inflammatory Response"]]
  [1] "A2m"       "Acvr1"     "Adam8"     "Adora1"    "Adra2a"    "Ahsg"      "Alox5ap"   "Ano6"      "Apcs"      "Apoa2"     "Apol8"     "Ash1l"    
 [13] "Ass1"      "B4galt1"   "Btk"       "C2cd4a"    "C2cd4b"    "C3"        "Ccr7"      "Cd163"     "Cd6"       "Cebpb"     "Cnr1"      "Creb3l3"  
 [25] "Crp"       "Ctnnbip1"  "Dnase1"    "Dnase1l3"  "Ednrb"     "Eif2ak1"   "Elane"     "Epo"       "Ext1"      "F12"       "F2"        "F3"       
 [37] "F8"        "Fcgr1"     "Fcgr2b"    "Fcgr4"     "Ffar2"     "Ffar3"     "Fn1"       "Fut7"      "Gata3"     "Gstp1"     "Gstp2"     "Hamp"     
 [49] "H2-T23"    "Hp"        "Ighg1"     "Ighg2b"    "Ighg2c"    "Ighg3"     "Il1a"      "Il1b"      "Il20rb"    "Il22b"     "Il31ra"    "Il4"      
 [61] "Il6"       "Il6ra"     "Il6st"     "Ins2"      "Itih4"     "Kl"        "Klkb1"     "Lbp"       "Mbl2"      "Mrgprb1"   "Mrgprb2"   "Mrgprx1"  
 [73] "Mylk3"     "Nlrp3"     "Nlrp6"     "Npy"       "Npy5r"     "Nupr1"     "Ogg1"      "Oprm1"     "Orm1"      "Orm2"      "Osm"       "Osmr"     
 [85] "Park7"     "Pik3cg"    "Pla2g2d"   "Plscr1"    "Prcp"      "Ptger3"    "Ptges"     "Ptgs2"     "Reg3b"     "Reg3g"     "Rhbdd3"    "S100a8"   
 [97] "Saa1"      "Saa2"      "Saa4"      "Selenos"   "Serpina1b" "Serpina1e" "Serpina3c" "Serpina3k" "Serpina3m" "Serpina3n" "Serpinf2"  "Sigirr"   
[109] "Tacr1"     "Tfr2"      "Tfrc"      "Tnf"       "Tnfrsf11a" "Tnfsf11"   "Tnfsf4"    "Trem1"     "Trpv1"     "Ugt1a1"    "Vcam1"     "Vnn1"     
[121] "Zp3"   
```{r inflam pathways}
pathways.show <- c("CCL", "CXCL", "COMPLEMENT", "CD86", "TNF", "TGFB")
# day 2
object.list <- object.list2
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))

# day 7
object.list <- object.list7
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "all", signaling = pathways.show, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "all", signaling = pathways.show, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```

# D0 to D2 to D7 Dotplots
workflow: compute the signaling probabilities for each group, then make a new matrix with just that data extracted and plot it in ggplot

original code
```{r}
# netAnalysis_river(cellchat_NP7, pattern = "outgoing")
# PBS is squares
h1 = netAnalysis_dot(cellchat_NP2, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 21)
h2 = netAnalysis_dot(cellchat_NP7, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 21)
h3 = netAnalysis_dot(cellchat_PBS2, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 22)
h4 = netAnalysis_dot(cellchat_PBS7, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 22)
h5 = netAnalysis_dot(cellchat_D0, pattern = "outgoing", group.show = c("Mon", "Neut"), shape = 22)

patchwork::wrap_plots(h3, h1, nrow=2)
patchwork::wrap_plots(h4,h2, nrow=2)

# dev.off()
```

plot from source code
```{r}
# plotting vars
shape = 21
dot.size = c(1,3)
dot.alpha = 1
main.title = NULL
font.size = 10
font.size.title = 12
color.use = NULL
cutoff = NULL
```

## Monocytes PBS

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% filter(CellGroup == "Mon")
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
```


```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Monocyte PBS Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08)))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Mon_D0_2_7_PBS_Dotplot.pdf", 
    width = 16, height = 3
    )
gg
dev.off()
```



workflow: compute the signaling probabilites for each group, then make a new matrix with just that data extracted and plot it in ggplot

original code
## Monocytes NP


```{r}
table <- h1[["data"]] %>% mutate(id = "D2_NP") %>% filter(CellGroup == "Mon")
new.table <- h2[["data"]] %>% mutate(id = "D7_NP") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
```


```{r}
gg2 <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Monocyte NP Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.1)))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg2 <- gg2 + guides(colour=FALSE) + guides(fill=FALSE)
  gg2 <- gg2 + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg2
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Mon_D0_2_7_NP_Dotplot.pdf", 
    width = 16, height = 3
    )
gg2
dev.off()
```

```{r}

```


## Monocytes All
```{r}
# table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% filter(CellGroup == "Mon")
# new.table <- h4[["data"]] %>% mutate(id = "D7_PBS") %>% filter(CellGroup == "Mon")
# table <- bind_rows(table, new.table)
# new.table <- h5[["data"]] %>% mutate(id = "D0") %>% filter(CellGroup == "Mon")
# table <- bind_rows(table, new.table)
# new.table <- h1[["data"]] %>% mutate(id = "D2_NP") %>% filter(CellGroup == "Mon")
# table <- bind_rows(table, new.table)
# new.table <- h2[["data"]] %>% mutate(id = "D7_NP") %>% filter(CellGroup == "Mon")
# table <- bind_rows(table, new.table)
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == "Mon")
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == "Mon")
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Monocyte Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Mon_D0_2_7_All_Dotplot.pdf", 
    width = 16, height = 3
    )
gg
dev.off()
```

## Neutrophils PBS

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% filter(CellGroup == "Neut")
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
```


```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Neutrophil PBS Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08)))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Neut_D0_2_7_PBS_Dotplot.pdf", 
    width = 16, height = 3
    )
gg
dev.off()
```



workflow: compute the signaling probabilites for each group, then make a new matrix with just that data extracted and plot it in ggplot

original code
## Neutrophils NP


```{r}
table <- h1[["data"]] %>% mutate(id = "D2_NP") %>% filter(CellGroup == "Neut")
new.table <- h2[["data"]] %>% mutate(id = "D7_NP") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
```


```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Neutrophil NP Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.1)))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```
```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Neut_D0_2_7_NP_Dotplot.pdf", 
    width = 16, height = 3
    )
gg
dev.off()
```

## Neutrophils All
```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == "Neut")
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == "Neut")
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Neutrophil Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(
      strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
          )
       )
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = "~/SCIFlexSeq/Figures/Cellchat/Neut_D0_2_7_All_Dotplot.pdf", 
    width = 16, height = 3
    )
gg
dev.off()
```

# Dot Plot of Communcation Networks
```{r}
netAnalysis_dot(cellchat_PBS2, pattern = "outgoing", group.show = c("Fib", "Stro"))

```
workflow: compute the signaling probabilities for each group, then make a new matrix with just that data extracted and plot it in ggplot

original code
```{r}
# netAnalysis_river(cellchat_NP7, pattern = "outgoing")
# PBS is squares
h1 = netAnalysis_dot(cellchat_NP2, pattern = "outgoing", group.show = c("Endo", "Micro", "Stro", "Fib", "Peri"), shape = 21)
h2 = netAnalysis_dot(cellchat_NP7, pattern = "outgoing", group.show = c("Endo", "Micro", "Stro", "Fib", "Peri"), shape = 21)
h3 = netAnalysis_dot(cellchat_PBS2, pattern = "outgoing", group.show = c("Endo", "Micro", "Stro", "Fib", "Peri"), shape = 22)
h4 = netAnalysis_dot(cellchat_PBS7, pattern = "outgoing", group.show = c("Endo", "Micro", "Stro", "Fib", "Peri"), shape = 22)
h5 = netAnalysis_dot(cellchat_D0, pattern = "outgoing", group.show = c("Endo", "Micro", "Stro", "Fib", "Peri"), shape = 22)

patchwork::wrap_plots(h3, h1, nrow=2)
patchwork::wrap_plots(h4,h2, nrow=2)

# dev.off()
```

plot from source code
```{r}
# plotting vars
shape = 21
dot.size = c(1,3)
dot.alpha = 1
main.title = NULL
font.size = 10
font.size.title = 12
color.use = NULL
cutoff = NULL
```


## Endothelial All
```{r}
# set cell type 
cell = "Endo"
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))


```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Endothelial Outgoing Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = paste0("~/SCIFlexSeq/Figures/Cellchat/", cell, "_D0_2_7_AllPatways_Outgoing_Dotplot.pdf"), 
    width = 16, height = 3
    )
gg
dev.off()
```

## Microglia 

```{r}
# set cell type 
cell = "Micro"
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Microglia Outgoing Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = paste0("~/SCIFlexSeq/Figures/Cellchat/", cell, "_D0_2_7_AllPatways_Outgoing_Dotplot.pdf"), 
    width = 16, height = 3
    )
gg
dev.off()
```

## Stromal

```{r}
# set cell type 
cell = c("Stro")
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Stromal Outgoing Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = paste0("~/SCIFlexSeq/Figures/Cellchat/", cell, "_D0_2_7_AllPatways_Outgoing_Dotplot.pdf"), 
    width = 16, height = 3
    )
gg
dev.off()
```

## Fibroblast

```{r}
# set cell type 
cell = c("Fib")
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Fibroblast Outgoing Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = paste0("~/SCIFlexSeq/Figures/Cellchat/", cell, "_D0_2_7_AllPatways_Outgoing_Dotplot.pdf"), 
    width = 16, height = 3
    )
gg
dev.off()
```

## Pericyte

```{r}
# set cell type 
cell = c("Peri")
```

```{r}
table <- h3[["data"]] %>% mutate(id = "D2_PBS") %>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
new.table <- h4[["data"]] %>% mutate(id = "D7_PBS")%>% mutate(treat = "PBS") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h5[["data"]] %>% mutate(id = "D0")%>% mutate(treat = "Naive") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h1[["data"]] %>% mutate(id = "D2_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table)
new.table <- h2[["data"]] %>% mutate(id = "D7_NP")%>% mutate(treat = "NP") %>% filter(CellGroup == cell)
table <- bind_rows(table, new.table) %>% filter(!is.na(Contribution))
```

```{r}
gg <- ggplot(data = table, aes(x = Signaling, y = id)) +
    geom_point(aes(size =  Contribution, fill = id, colour = id), shape = 21) +
    facet_wrap(~treat, nrow=3, scales = "free_y",  strip.position = "right") +
    scale_size_continuous(range = dot.size) +
    theme_linedraw() +
    scale_x_discrete(position = "bottom") +
    ggtitle("Pericyte Outgoing Signaling Pathways") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme(text = element_text(size = font.size),plot.title = element_text(size=font.size.title, face="plain"),
          axis.text.x = element_text(angle = 45, hjust=1),
          axis.text.y = element_text(angle = 0, hjust=1),
          axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    theme(axis.line.x = element_line(size = 0.25), axis.line.y = element_line(size = 0.25)) +
    theme(panel.grid.major = element_line(colour="grey90", size = (0.08))) +
    theme(strip.text = element_text(colour = "black"),
        strip.background = element_rect(
          fill = "grey90",
         color = "black",
          linewidth = 1
  ))
  # gg <- gg + scale_y_discrete(limits = rev(levels(data3$CellGroup)))
  # gg <- gg + scale_fill_manual(values = ggplot2::alpha(color.use, alpha = dot.alpha), drop = FALSE, na.value = "white")
  # gg <- gg + scale_colour_manual(values = color.use, drop = FALSE, na.value = "white")
  gg <- gg + guides(colour=FALSE) + guides(fill=FALSE)
  gg <- gg + theme(legend.title = element_text(size = 10), legend.text = element_text(size = 8))
  gg
```

```{r}
pdf(file = paste0("~/SCIFlexSeq/Figures/Cellchat/", cell, "_D0_2_7_AllPatways_Outgoing_Dotplot.pdf"), 
    width = 16, height = 3
    )
gg
dev.off()
```

# Circle Plots
```{r}
# Chord diagram
par(mfrow = c(1,2), xpd=TRUE)
netVisual_chord_gene(object.list[[2]], sources.use = 4, targets.use = c(5:11), slot.name = 'net', net = net.up, lab.cex = 0.8, small.gap = 3.5, title.name = paste0("Up-regulated signaling in ", names(object.list)[2]))
netVisual_chord_gene(object.list[[1]], sources.use = 4, targets.use = c(5:11), slot.name = 'net', net = net.down, lab.cex = 0.8, small.gap = 3.5, title.name = paste0("Down-regulated signaling in ", names(object.list)[2]))
#> You may try the function `netVisual_chord_cell` for visualizing individual signaling pathway
```



# Simplified Plots
The differential number of interactions or interaction strength in the cell-cell communication network between two datasets can be visualized using circle plot, where red
 (or blue) colored edges represent increased
 (or decreased) signaling in the second dataset compared to the first one.
```{r}
gg1 <- netVisual_heatmap(cellchat7)
#> Do heatmap based on a merged object
gg2 <- netVisual_heatmap(cellchat7, measure = "weight")
#> Do heatmap based on a merged object
gg1 + gg2
```

```{r}
png(file = "~/SCIFlexSeq/Figures/Cellchat/D7_PBS_NP_Heatmap.png", 
    width = 8, 
    height = 5, unit= "in", bg = "transparent", res = 300 )
gg1 + gg2
dev.off()
```

```{r}
gg1 <- netVisual_heatmap(cellchat2)
#> Do heatmap based on a merged object
gg2 <- netVisual_heatmap(cellchat2, measure = "weight")
#> Do heatmap based on a merged object
gg1 + gg2
```

```{r}
png(file = "~/SCIFlexSeq/Figures/Cellchat/D2_PBS_NP_Heatmap.png", 
    width = 8, 
    height = 5, unit = "in", res = 300, bg = "transparent")
gg1 + gg2
dev.off()
```


```{r}
par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat7, weight.scale = T)
netVisual_diffInteraction(cellchat7, weight.scale = T, measure = "weight")

weight.max <- getMaxWeight(object.list7, attribute = c("idents","count"))
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list7)) {
  netVisual_circle(object.list7[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list7)[i]))
}
ggsave(file = "~/SCIFlexSeq/Figures/Cellchat/D7_interaction_condition.png")

```

```{r}
par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat2, weight.scale = T)
netVisual_diffInteraction(cellchat2, weight.scale = T, measure = "weight")

weight.max <- getMaxWeight(object.list2, attribute = c("idents","count"))
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list2)) {
  netVisual_circle(object.list2[[i]]@net$count, weight.scale = T, label.edge= F, edge.weight.max = weight.max[2], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list2)[i]))
}
```

## Simplified Names
Simplify names of D7 cellchat objects, as 'cellchat' and 'object.list7'
```{r}
group.cellType <- c("Stro", "Micro", "Mon", "Endo", "Micro", "Leuko", "Glial", "Mon", "Glial", "Glial", "Stro", "Glial", "Stro", "Neut", "Stro")
group.cellType <- factor(group.cellType, levels = c("Mon", "Neut", "Stro", "Endo", "Micro", "Glial", "Leuko"))
object.list7 <- lapply(object.list7, function(x) {mergeInteractions(x, group.cellType)})
cellchat <- mergeCellChat(object.list7, add.names = names(object.list7))


weight.max <- getMaxWeight(object.list7, slot.name = c("idents", "net", "net"), attribute = c("idents","count", "count.merged"))
par(mfrow = c(1,2), xpd=TRUE)
for (i in 1:length(object.list7)) {
  netVisual_circle(object.list7[[i]]@net$count.merged, weight.scale = T, label.edge= T, edge.weight.max = weight.max[3], edge.width.max = 12, title.name = paste0("Number of interactions - ", names(object.list7)[i]))
}
```

differential interactions
```{r}
par(mfrow = c(1,2), xpd=TRUE)
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "count.merged", label.edge = T)
netVisual_diffInteraction(cellchat, weight.scale = T, measure = "weight.merged", label.edge = T)
```

```{r}
library(ComplexHeatmap)
object.list <- object.list7

i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 6)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 6)
draw(ht1 + ht2, ht_gap = unit(0.5, "cm"))
```

```{r}
mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

